Speculation in foreign exchange: noise or information?

### 2015-10-20 Knitr/ElsPaper.tex
Created the ElsPaper.tex Need to check the key words and the way that the author is presented.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed and noise.      
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices. 
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.

@@

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

@@

The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.

\section{Informed and uninformed speculation}
Speculation plays an important part in several areas of financial economics.  However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making. A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.   Speculation and trading can be categorised as being informed or uninformed. Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.

@@

Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}

@@

\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.  However, a large proportion of uninformed traders in a market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will drive prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse. This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation} 
 where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}

@@

\section{Measuring speculation and the event study}
There are two steps: identify speculation and assess what happens when it becomes extreme.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium and the following price performance should be a random walk.

The \citet[pp. 14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.  It is summarised by Equation \ref{eq:NP2}. There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse. The following paragraphs show how these can be measured. \footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted. As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.} This reversal is the Friedman-effect discussed above.

The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.  In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put. The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

@@

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  The Euro data runs from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}  
An \emph{event study} is used to assess the effect of extreme speculation on future returns.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study. The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.  Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.

@@

  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).  The windows were chosen to correspond to one day and one week of trading.  However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, makes it clear that altering the size of the event window does not change the findings to any significant degree.

@@

\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance means that more than 95\% of 1000 random bootstraps from the events are above or below zero respectively.   

@@

That sentiment drives price appears evident from the covariance of speculative sentiment with exchange rate prices. This finding is consistent with the evidence of previous research from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}. If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, this suggests that there is information that the equilibrium is very different from the current price.

What is interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.  Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation rather than a reversal that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the $99^{th}$ percentile is used to measure the extreme, there is even more ambiguity.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant. Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes. Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.  
\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES5}
\end{figure}

@@

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16a}
\caption{Event Study:  Extreme (Low - $10^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES2}
\end{figure}

While there is some evidence that speculation is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to news.  There is a gradual absorption of information and a gradual evolution of prices.  This is consistent with the evidence of serial correlation in asset prices that has been identified.  See, for example, \citet{Jagadeesh}.

Two other tests have been carried out to understand more about speculative extremes: the distribution of exchange rate returns after the extreme are compared to the distribution of returns for the whole sample; the level of speculative sentiment is identified when large positive and negative returns are recorded. The distribution data is provided in Table \ref{tabref:D}.  The most striking feature is the broader standard deviation that is recorded. 
% Add the table with the descriptive statistics for returns that follow an extreme compared to the whole sample. 
\begin{table}[h]
\begin{threeparttable}
\begin{tabular}{rrrrrrr}
& \multicolumn{4}{c}{Extreme Distribution}&\multicolumn{2}{c}{Regular Distribution}\\
& \multicolumn{2}{c}{High}&\multicolumn{2}{c}{Low}&
\multicolumn{2}{l}{}\\
 & Mean & Std & Mean & Std & Mean & Std \\
\hline
AUD & 0.0125 & 0.8312 & -0.1711 & 1.0588 & -0.0174 & 0.7241 \\ 
  EURGBP & 0.0696 & 0.7582 & 0.0996 & 0.9028 & -0.0187 & 0.5542 \\ 
  EURSEK & 0.0033 & 0.6572 & 0.1373 & 0.6059 & 0.0011 & 0.4384 \\ 
  GBP & 0.1909 & 0.8861 & 0.0312 & 0.9314 & 0.0025 & 0.4892 \\ 
  CHF & -0.0565 & 0.9577 & -0.1125 & 1.3452 & 0.0120 & 0.6819 \\ 
  EURCHF & 0.0690 & 0.4198 & 0.0194 & 0.5445 & -0.0039 & 0.2769 \\ 
  EURJPY & -0.1103 & 1.2729 & -0.1647 & 1.5424 & -0.0072 & 0.8416 \\ 
  EUR & 0.1586 & 1.0942 & -0.0933 & 0.9411 & -0.0172 & 0.6338 \\ 
  CAD & -0.0489 & 0.8006 & -0.0571 & 0.5236 & 0.0084 & 0.3632 \\ 
  JPY & 0.1341 & 1.0090 & 0.0946 & 1.3921 & 0.0076 & 0.8170 \\ 
   \hline
\caption{Comparison of extreme and regular return distribution}
\label{tabref:D}
\end{tabular}
\begin{tablenotes}
\small 
\item This table shows a comparison of the mean return for the whole of the sample compared to the mean return for the day after an extreme has been reached.  It is clear that there is no significant different between the mean returns, arguing against the idea that speculation is uninformed. However, the standard deviation of the returns is much broader once there has been extreme speculation. This may suggest increased risk or increased uncertainty.  
\end{tablenotes}
\end{threeparttable}  
\end{table}

>>>>

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\begin{frontmatter}
\title{Speculation in foreign exchange: noise or information?}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}

\begin{abstract}
If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.      
\end{abstract}
\begin{keyword}
JEL classification: G12, G14, G15; information, noise-trading, event studies.
\end{keyword}

\end{frontmatter}
\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices. 
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.

@@

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after extremes.

@@

The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the event study method; Section Four reviews the evidence; Section Five concludes.

\section{Informed and uninformed speculation}
Speculation plays an important part in several areas of financial economics.  However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making. A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.   Speculation and trading can be categorised as being informed or uninformed. Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.  See \citet{Ramiah201589} for an overview of the noise-trader literature.

@@

Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}. The aim is to determine not the underlying value but what others perceive to be the underlying value.}

@@

\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.  However, a large proportion of uninformed traders in a market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse. This can be called \emph{noise-trader risk}.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation} 
 where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}

@@

\section{Measuring speculation and the event study}
There are two steps: identify speculation and assess what happens when it becomes extreme.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation achieves a new equilibrium and the following price performance should be a random walk.

The \citet[pp. 14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.  It is summarised by Equation \ref{eq:NP2}. There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse. This reversal is the Friedman-effect discussed above. The following paragraphs show how these can be measured. \footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted. As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.}

The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.  In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put. The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  %A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

@@

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  The Euro data runs from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}

An \emph{event study} is used to assess the effect of extreme speculation on future returns.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study. The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.  Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.

@@

  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).  The windows were chosen to correspond to one day and one week of trading.  However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, and it becomes clear that altering the size of the event window does not change the findings to any significant degree.

@@

\item Where Hi means extreme high in the risk-reversal skew and Lo means extreme low in the risk reversal skew; extreme is calculated as being equal or above the $95^{th}$ or $99^{th}$ percentile or equal or below the $5^{th}$ or $1^{st}$ percentile respectively; CARW is the cumulative abnormal return for the whole window, before and after the extreme event, and CARA is the cumulative abnormal return for the period after the event, which is the event day and the window; abnormal return is anything that is different from zero; the asterisk denotes significantly different from zero, where statistical significance is more than 95\% of 1000 means calculated from random bootstrap samples from the events are above or below zero respectively.   

@@

That sentiment drives price appears evident from the covariance of speculative sentiment with exchange rate prices. This finding is consistent with the evidence of previous research that has used order-flow or sentiment from the likes of \citet{Evans2002Order}, \citet{FuturesSanders} and \citet{FuturesWang}. If speculation is not informed, the extreme is caused by herding or institutional features; if speculation is informed, it is facilitating the absorption of this information into the price.

Does knowledge about the extreme at event time $t = 0$ provide information about the future returns?  If this speculation is noise, informed traders, such as real money or long-term fundamental accounts, should take advantage of the space that has been created between the price and the fundamental value and a reversal should be seen.  Whether this is the case is recorded in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the $5^{th}$ and $95^{th}$ percentiles to identify extremely negative and extremely positive sentiment respectively, ten show a continuation (rather than a reversal) that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal - EURCHF in each case.  For the more extreme event when the $99^{th}$ percentile is used, there is even less clarity.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant. Exchange rate prices are as likely to go up as down after an extreme and there does not appear to be any information about the future from knowledge of these extremes. Extremes seem to be followed by a random walk as would be the case if the extreme were the result of information being absorbed into the price.  
\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES5}
\end{figure}

@@

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16a}
\caption{Event Study: Extreme (Low $10{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or below the $10^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES2}
\end{figure}

While there is some evidence that speculation is informed, it is clear from figures \ref{fig:ES1} and \ref{fig:ES2} that exchange rates do not react instantly to news.  There is a gradual absorption of information and a gradual evolution of prices.  Serial correlation in asset prices is a consistent observation.  See \citet{Jagadeesh}. In this case the serial correlation is associated with extreme speculation that appears to be informed.

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study: Extreme (High $99{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES1}
\end{figure}

Table \ref{tabref:SP1} shows the results from the event studies that were carried out using speculative positions as reported to US regulators.  The table works in a similar way to Table \ref{tabref:RR1}.  The first column is the exchange rate, defined as units per US dollar.  The table is then split into two parts.  The first part uses the S1 measure, the net non-commercial (speculative) positions relative to total non-commercial (speculative) positions; the second is S2 as the net non-commercial position relative to open interest or the total outstanding open positions.  Therefore, the first captures the balance of speculative positions while the second captures the balance of speculative positions relative to the whole market.  If speculators become more prevalent or dominant in the market, this will show up in S2 but not S1.  S1 is more sensitive to changes in sentiment and more volatile. Each of these sections is broken into the results for a 2 week window and a 4 week window.  There is a row for extreme high and extreme low which represents extreme long positions or extreme short positions; the first  column identifies the number of cases; CARW is the cumulative abnormal return for the whole event window; CARA is the cumulative abnormal return for the period after the event to the end of the window.
